---
title: "Hands-on_Ex02"
author: "Xi Zixun"
---

# 1.0 Set up

### 1.1 Checking Required Packages

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, tidyverse, hrbrthemes) 
```

### 1.2 Data Loading

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## 2.0 Beyond ggplot2 Annotation : ggrepel

::: panel-tabset
## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS,
       y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label(aes(label = ID),
             hjust= .5,
             vjust= .5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English Score versus Maths Scores For Primary 3")
```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS,
       y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label(aes(label = ID),
             hjust= .5,
             vjust= .5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English Score versus Maths Scores For Primary 3")
```
:::

### 2.0.1 Working With Repel

::: panel-tabset
## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS,
       y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label_repel(aes(label = ID),
             hjust= .5,
             vjust= .5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English Score versus Maths Scores For Primary 3")
```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS,
       y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label_repel(aes(label = ID),
             hjust= .5,
             vjust= .5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English Score versus Maths Scores For Primary 3")
```
:::

## 2.1 Beyond ggplot2 themes

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  theme_gray() +
  ggtitle("Distributon of Math Scores")
```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  theme_gray() +
  ggtitle("Distributon of Math Scores")
```
:::

### 2.1.1 Working With ggtheme Package

::: panel-tabset
## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins=20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  ggtitle('Distribution of Math Scores') +
  theme_economist()
```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins=20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  ggtitle('Distribution of Math Scores') +
  theme_economist()
```
:::
It also provide some extra geoms and scales for '**gglot2**'.

### 2.1.2 Working with hrbrtheems Package

hrbrthemes package provides a base theme that focus on typographic elements, including where various labels are placed as well as the fonts that are used.

::: panel-tabset
## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()

```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
:::

The second goal centers around productivity for a production workflow. In face, this '**production workflow**' is the context for where the elements of **hrbrthemes** could be used.

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') + 
  ggtitle('Distrbution of Math Scores') +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = 'Y')

```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') + 
  ggtitle('Distrbution of Math Scores') +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = 'Y')

```

:::

## 2.2 Beyond Single Graph

It is not unusual that multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graph. In this section we will learn how to create composite plot by combining multiple graphs. First, let us create three statistical graphics by using the code chunk below

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
p1 <- ggplot(data = exam_data,
             aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  coord_cartesian(xlim = c(0,100)) +
  ggtitle('Distribution of Math Scores')
p1

```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
p1 <- ggplot(data = exam_data,
             aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  coord_cartesian(xlim = c(0,100)) +
  ggtitle('Distribution of Math Scores')
p1

```
:::

Next

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
p2 <- ggplot(data = exam_data,
             aes(x = ENGLISH)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  coord_cartesian(xlim = c(0,100)) +
  ggtitle('Distribution of English Scores')
p2

```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
p2 <- ggplot(data = exam_data,
             aes(x = ENGLISH)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = 'grey25',
                 fill = 'grey90') +
  coord_cartesian(xlim = c(0,100)) +
  ggtitle('Distribution of English Scores')
p2

```
:::

Lastly, draw a **Scatter Plot** for English vs Math Score as shown below 

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
p3 <- ggplot(data = exam_data,
             aes(x = MATHS,
                 y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm,
              size = 0.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle(' English VS Math Scores - Primary 3')
p3

```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
p3 <- ggplot(data = exam_data,
             aes(x = MATHS,
                 y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm,
              size = 0.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle(' English VS Math Scores - Primary 3')
p3

```
:::

### 2.2.1 Creating Composite Graphics: Pathwork Method

There are several ggplot2 extension's functions support the needs to prepare composite figure by combining several graphs such as grid_arrange() of **gridExtra** package and plot_grid() of **cowplot** package.  In this section we are going to shared about ggplot2 extension call **patchwork** which is specially designed for combining separate ggplot2 graphs into a single figure.

Patchwork packages has very simple syntax where we can create layouts super easily. Here are the general syntax that combines:

  · Two-Column Layout using Plus Sign `+` \
  · Parenthesis() to create a subplot group. \
  · Two-Row Layout using the Division Sing `/` \
  
  
### 2.2.2 Combining Two ggplot2 graphs

Figure in the tabset below wos a compsite of two histogram created using patchwork. \
Note how simple the syntax used to create the plot!

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
p1 + p2
```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
p1 + p2
```

:::


### 2.2.3 Combining Three ggplot2 graphs

We can plot more complex composite by using appropriate operators. For example, the composite figure below is plotted by using:

  · `/` operator to stack two ggplot2 graphs \
  · `|` operator to place the plots beside each other \
  · `()` operator to define the sequence of the plotting \
  
  
::: panel-tabset
## Plot
```{r}
#| echo: false
#| message: false
#| warning: false
(p1 / p2) | p3
```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
(p1 / p2) | p3
```
:::

### 2.2.4 Creating a composite Figure with tag

In order to identify subplot in text, **patchwork** also provides auto-tagging capabilities as shown in the figure below.

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
((p1 / p2) | p3) +
  plot_annotation(tag_levels = 'I')

```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
((p1 / p2) | p3) +
  plot_annotation(tag_levels = 'I')

```

:::
  
### 2.2.5 Creating Figure with insert
Beside providing functions to place plots next to each other based on the provided layouts. \
With ```inset_element()``` of **patchwork**, we can place on or several plots or graphic elements freely on top or bloe another plot.

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
p3 + inset_element(p2,
                   left = 0.02,
                   bottom = 0.7,
                   right = 0.5,
                   top = 1)
```


## Code

```{r}
#| eval: false
#| message: false
#| warning: false
p3 + inset_element(p2,
                   left = 0.02,
                   bottom = 0.7,
                   right = 0.5,
                   top = 1)
```

:::

### 2.2.6 Creating a composite figure by using patchwork & ggtheme
Figure below is created by combining patchwork and theme_economist() of ggthemes package discussed earlier.

::: panel-tabset

## Plot

```{r}
#| echo: false
#| message: false
#| warning: false
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()

```

## Code

```{r}
#| eval: false
#| message: false
#| warning: false
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()

```

:::

## 2.3 Reference

  · [Patchwork Rpackage goes nerd viral](https://www.littlemissdata.com/blog/patchwork) \
  · [ggrepel](https://ggrepel.slowkow.com/) \
  · [ggthemes](https://ggplot2.tidyverse.org/reference/ggtheme.html) \
  · [hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/) \
  · [ggplot tips: Arranging Plots](https://albert-rapp.de/post/2021-10-28-extend-plot-variety/) \
  · [ggplot2 Theme Elements Demonstration](https://henrywang.nl/ggplot2-theme-elements-demonstration/) \
  · [ggplot2 Theme Elements Reference Sheet](https://isabella-b.com/blog/ggplot2-theme-elements-reference/) \